---
title: "Reproducible reaserch, Course Project 1"
author: "Mario Israel Hernández"
date: "8 de septiembre de 2020"
output: html_document
---
First we need to read the file with the observations and transform them into a suitable dataframe.
```{r}
raw_data <- read.csv("activity.csv")
dates <- as.Date(raw_data$date)
data1 <- cbind(raw_data, dates)
NiceData <- data1[,c(1,3,4)]
```

So, now that we have the nice tidy dataframe we were looking for, we can start manipulating it.
```{r}
library(ggplot2)

weekday <- weekdays(NiceData$dates)
Fulldata <- cbind(NiceData,weekday)

summary(Fulldata)
```
So, now that we have add a tag for every day of te week, we can calculate the mean total number of steps taken per day.
```{r}
activity_total_steps <- with(Fulldata, aggregate(steps, by = list(dates), FUN = sum, na.rm = TRUE))
names(activity_total_steps) <- c("date", "steps")
hist(activity_total_steps$steps, main = "Total number of steps taken per day", xlab = "Total steps taken per day", col = "darkblue", ylim = c(0,20), breaks = seq(0,25000, by=2500))
##Mean
mean(activity_total_steps$steps)
##Median
median(activity_total_steps$steps)
```
Now we calculate the average of daily activity.
```{r}
average_daily_activity <- aggregate(Fulldata$steps, by=list(Fulldata$interval), FUN=mean, na.rm=TRUE)
names(average_daily_activity) <- c("interval", "mean")
plot(average_daily_activity$interval, average_daily_activity$mean, type = "l", col="darkblue", lwd = 2, xlab="Interval", ylab="Average number of steps", main="Average number of steps per intervals")
```
And to answer the interval max question:
```{r}
average_daily_activity[which.max(average_daily_activity$mean), ]$interval
```
To do the input missing values first we need to know how many of them are missing:
```{r}
sum(is.na(Fulldata$steps))
```
Now we calculate the median of the day.
```{r}
imputed_steps <- average_daily_activity$mean[match(Fulldata$interval, average_daily_activity$interval)]
```
And generate a new data frame.
```{r}
activity_imputed <- transform(Fulldata, steps = ifelse(is.na(Fulldata$steps), yes = imputed_steps, no = Fulldata$steps))
total_steps_imputed <- aggregate(steps ~ dates, activity_imputed, sum)
names(total_steps_imputed) <- c("date", "daily_steps")
```
Aaand plot!
```{r}
hist(total_steps_imputed$daily_steps, col = "darkblue", xlab = "Total steps per day", ylim = c(0,30), main = "Total number of steps taken each day", breaks = seq(0,25000,by=2500))
```
The mean
```{r}
mean(total_steps_imputed$daily_steps)
```
and median
```{r}
median(total_steps_imputed$daily_steps)
```
To determine the difference between the weekdays and weekends, we first have to tag them properly:
```{r}
Fulldata$date <- as.Date(strptime(Fulldata$date, format="%Y-%m-%d"))
Fulldata$datetype <- sapply(Fulldata$date, function(x) {
        if (weekdays(x) == "Sábado" | weekdays(x) =="Domingo") 
                {y <- "Weekend"} else 
                {y <- "Weekday"}
                y
        })
```
And then plot them separetly with a time series on the x axis.
```{r}
activity_by_date <- aggregate(steps~interval + datetype, Fulldata, mean, na.rm = TRUE)
plot<- ggplot(activity_by_date, aes(x = interval , y = steps, color = datetype)) +
       geom_line() +
       labs(title = "Average daily steps by type of date", x = "Interval", y = "Average number of steps") +
       facet_wrap(~datetype, ncol = 1, nrow=2)
print(plot)
```

